\documentclass{article}
\usepackage{amsmath,amsthm,amsfonts}                        % AMS Math
\usepackage{thmtools}                                       % Theorem Tools
\usepackage{bm}        
\usepackage{tabto}                                          % Bold Math

\title{MATH1081 notes}
\author{Nira (z5417727)}
\date{September 17th, 2022}
\begin{document}
\maketitle 
\tableofcontents
\newpage
\section{Topic 1}
\subsection{Introduction}
1. addition, multiplication, division and subtraction
\newline
2. Mainly dealing with finite sets

\newpage
\subsection{Sets and subsets}
\boxed{ $\text{A set is a well defined collection of distinct objects}$ }
\newline
\newline
Example: $S = \{1, a, 3\}, A =\{\Pi, 1\}.$
\newline
1. $e \notin A$; it is not in A
\newline
2. For example, if A is a set of all integers; $\{ \text{all even integers} \}$
    = $\{n \in \mathbb{R} | \text{n is even}\}$.
\newline
3. We can remove superfluos items (elements that occur more than one).
$A = \{ 1, 2, 3, 3 \}$ where $3$ can be removed.
\newline
\newline
\newline
Example:
\newline
$A = \{ 1, 2, 3 \},
B = \{2, 3, 1 \}, 
C = \{1, 2, 3, 3 \},
D = \{ 1 , 3 \}.$
\newline
Here, D is a proper subset of A, B, C; A, B, C are supersets of D.
\newline
\boxed {$$\subseteq: \text{Subset (proper subset)}, $$
$$ \supseteq: \text{Superset}.$$}
\newline
\newline
1. To prove if a set is a proper subset; do the following:
\newline
For example, if $D \in A$, then check if $e \in D$
\newline
If $e \in D,$ then $e \in A$.
Thus, it would be a proper subset (here, e is just an element).
\newline
\newline
2. To prove that two sets are equal;
\newline
For example, if A = B, prove:
\newline
i) $A \subseteq B$; if an element is in A, then the element is in B.
\newline
ii) $B \subseteq A$; if an element is in B, then the element is in A.
\newpage
\subsection{Power Sets and Stability}
Subsets of $A = \{1, 2, 3\} $:
\newline
1. Could throw everything out to get empty set $\Phi$,
\newline
2. One element each: $\{ 1 \}, \{ 2 \}, \{ 3 \}$,
\newline
3. Two elements: $\{ 1, 2 \}, \{ 2, 3 \}, \{ 1, 3 \}$,
\newline
4. Set itself: $A$.
\newline
The set containing 1, 2, 3, 4 is called the powerset of A.
\newline
\newline
Given $
A = \{1, 2, 3\}, B = \{1, 2, 3, 3\}, C = \{1, 3\}, D = \{1, 3\}$, where \newline
$A = B$, $C \subseteq A, B$ and $D \not \subseteq A, B, C$.
\newline
1. size of A = 3, B = 3, C = 2, D = 2.
\newline
[Exercise with A  = {0, 1, {0, 1}}, B done in word].
\newpage
\subsection{Set Operations}
Boolean Operators ("not" operation in programming): 
\newline
1. Complement:
\newline
Let there be a set A in U (A: all of the people in the video, U: universal set of everyone in the world, $A^c =$ complement of A).
\newline
\newline
\boxed { $$A^c = \{ x \in U | x \not \in A \}.$$}
\newline
\newline
2. Intersecting ("and" operation in programming):
\newline
If there is $A, B$, intersecting, 
\newline
\newline
\boxed{ $$A \cap B = \{ x \in A | x \in B\}.$$}
\newline
\newline
3. Union ("or" operation in programming):
If there is $A, B$, A or B is:
\newline
\newline
\boxed{ $$A \cup B = \{ x \in U | x \in A \text { or } x \in B\}.$$}
\newline
\newline
4. Difference:
If there is $A, B$, intersecting, 
\newline
\newline
\boxed{ $$A - B = \{ x \in A | x \in B \}.$$}
\newline
\newline
[examples in word doc]
\newpage
\subsection{The Inclusion-Exclusion Principle}
[example in Word]
\newline
\boxed{ $$|A \cup B| = |A| + |B| - | A \cap B|$$.}
\newline
\newline
For three elements, 
\newline
\newline
\boxed{ $$|A \cup B \cup C| = |A| + |B| + |C| - | A \cap B| - |A \cap C| - |B \cap C| + |A \cap B \cap C|$$.}
\newline
\newline
[example in word]
\newpage
\subsection{Sets Proofs}
[proof question in word]
\newline
Hints for proofs:
\newline
1. To prove that $S \subseteq T$, we can assume that $x \in S$ and show that $x \in T$.
\newline
2. 
To prove that $S = T$, we can show that $S \subseteq T$ and $T \subseteq S$.
\newline
\newline
Scaffold: 
\newline
\boxed{ \text{Proof: Suppose that \dots 
\newline
\dots (proof)
\newline
we see that/ it follows \dots (conclusion) (end with shaded box to indicate end of proof.)}}
\newline
\newline
Note that the "Suppose that" part of the proof is usually whatever the if statment mentions.
\newline
\newline
For example, if the question is "Prove that if $A \cap B = A$, then $A \cup B = B$, then the proof starts like this:
\newline
\boxed{\underline{Proof}: \text{ Suppose that } A \cap B = A. }
\newline
For questions like "is this statement true", there are two ways to approach the question:
\newline
1. If the statement is true (if you think it is true), then prove it.
\newline
2. If the statement is false, then give a counter-example that proves it false. 
\newline
[examples in word]
\newpage
\subsection{Laws of Set Algebra}
\underline{ Laws of Set Algebra}
\newline
$$1\text{. } A \cap B = B \cap A: \text{ Commutative Law}.$$
$$2\text{. } A \cap \text{(} B \cap C \text{)} = \text{(} A \cap B \text{)} \cap C: \text{ Associative Law}.$$
$$3\text{. } A \cap \text{(} B \cap C \text{)} = \text{(} A \cap B \text{)} \cup \text{(} A \cap C \text{)}: \text{ Distributive Law}.$$
$$4\text{. } A \cap \text{(} A \cap B \text{)} = A: \text{ Absorption Law}.$$
$$5\text{. } A \cap U = U \cap A = A: \text{ Identity Law}.$$
$$6\text{. } A \cap A = A: \text{ Idempotent Law}.$$
$$7\text{. } (A^c)^c = A: \text{ Double Complement Law}.$$
$$8\text{. } A \cap \emptyset = \emptyset \cap A = \emptyset: \text{ Domination Law}.$$
$$9\text{. } A \cap A^c = \emptyset: \text{ Intersection with Complement Law}.$$
$$10\text{. } (A \cup B)^c = A^c \cap B^c: \text{ De Moirve's Law}.$$
The intersection can be swapped with the union to form another law (like, $A \cup B = B \cup A$ swapped as $A \cap B = B \cap A.$). Similarily, $U$ should be swapped with $\emptyset$ and vice versa.
\newline
[examples in word]
\newpage
\subsection{Generalised Set Operations}
Unions and Intersections; A saga:
$$1 \text{. } \cup_{i=1}^{n}A_i = A_1 \cup A_2 \cup \dots \cap A_n,$$
$$2 \text{. } \cap_{i=1}^{n} A_i = A_1 \cap A_2 \cap \dots \cap A_n.$$
Example:
\newline
\newline
$$A_k = {k, k + 1};$$
$$= \cup_{i=1}^{3}A_k = A_1 (\{1, 2\}) \cup A_2 (\{2, 3\}) \cup A_3 (\{3, 4\}),$$
$$= \{1, 2, 3, 4\}.$$
[example in word]
\newpage
\subsection{Russel's Paradox}
A set may contain another set as one of its elements.
\newline
This raises the possibility that a set may contain itself as an element.
\newline
\newline
\textbf{Problem: Try to let S be the set of all sets that are not elements of themselves, i.e.,}
$\mathbf{S = \{ A | A }\textbf{ is a set and } \mathbf{A \not\in A \}}.$
\newline
\textbf{Is S an element of itself?}
\newline
i)	If $S \in S$, then the definition of S implies that $S \not \in S$, a contradiction.
\newline
ii)	If $S \not \in S$, then the definition of S implies that $S \in S$, also a contradiction.
Hence neither $S \in S$ nor $S \not \in S$. This is Russell’s paradox.
\newpage
\subsection{Cartesian Product}
[example in word]
\newline
The Cartesian product of two sets A and B, denoted by $A \times B$, is the set of all ordered pairs from A to B:
\newline
\newline
\boxed{A \times B = \{(a,b)|a \in A \text{ and } b \in B\}}
\newline
\newline
If $|A| = m$ and $|B| = n$, then we have $|A \times B| = mn$.
\newline
\newline
Sets with more than 2 elements:
\newline
\newline
\textbf{Example:} $A = \{ a, b \}, B = \{ 1, 2, 3 \}.$
$$\text{Cartesian Product }(A \times B) = \{(a, 1), (a, 2), (a, 3), (b, 1), (b, 2), (b, 3)\}$$
(all of the ordered pairs -- combinations)
\newline
\newline
[example in word]
\newline
\newline
When X and Y are small finite sets, we can use an arrow diagram to represent a subset S of $X \times Y$ : we list the elements of X and the elements of Y , and then we draw an arrow from x to y for each pair $(x,y) \in S$.
\newpage
\subsection{Functions}
Example: Take 2 sets $X$ and $Y$, for which we have to find a function.
$$X = \{\text{all MATH 1081 students} \}, Y = \{ 0, 1, \dots , 84, 85, \dots, 100\}.$$
X: number of students; Y: marks from $0 - 100$.
\newline
Take function $f: X \to Y; \text{where } X \text{ is the domain and } Y \text{ is the co domain}.$
\newline
Ie, $f(x)$ = X's mark (Y).
\newline
Function $f: X \to Y$ satisfies $\{ (x, f(x)) | x \in X\} \subseteq X \times Y \text{ so that, for each } x \in X$;
\newline
1. $f(x)$ exists
\newline
2. $f(x)$ is unique
\newline
\newline
[example in word]
\newline
\newline
\boxed{\text{Note: be vary of the one-to-one function property lol}}
\newline
\newline
\underline{ Floor function and ceiling functions:}
\newline
\newline
1. Floor function (rounds down; smallest integer):
\newline
\newline
\boxed{\lfloor x \rfloor = \text{ max } \{ z \in Z | z \leq x\}.}
\newline
\newline
2. Ceiling function (rounds up; largest integer):
\newline
\newline
\boxed{\lceil x \rceil = \text{ min } \{ z \in Z | z \geq x\}.}
\newline
\newline
[example in word]
Domain/codomain: $\lfloor x \rfloor / \lceil x \rceil: \mathbb{R} \to \mathbb{Z}.$
\newline
\newline
Range($\lceil x \rceil$) = $\mathbb{Z}.$
\newline
\newline
[example in word]
\newpage
\subsection{Image and Inverse Image}
- The image of a set $A \subseteq X$ under a function $f : X \rightarrow Y \text{ is } f(A) = \{ y \in Y \text{ | } y = f(x) \text{ for some } x \in A\} = \{f(x)|x \in A\}$.
\newline
\newline
- The inverse image of a set $B \subseteq Y$ under a function $f : X \rightarrow Y is f^{-1}(B) = \{ x \in X |f(x) \in B \}$.
\newline
\newline
(image is just function values in the doman and inverse image is function values in range).
\newline
\newline
\boxed{\text{note: this is just function and inverse functions.}}
\newline
\newline
[example in word]
\newpage
\subsection{Injective, Surjective, Bijective}
Formal Definitions:
\newline
\newline
Recall that if f is a function from X to Y , then   for every $x \in X$, there is exactly one $y \in Y$ such that $f(x) = y$.
\newline
\newline
  1. We say that a function $f : X \rightarrow Y$ is injective or one-to-one if, for every $y \in Y$ , there is at most one $x \in X$ such that $f(x) = y$. 
  \newline
  Example: for all $x_1,x_2 \in X$, if $f(x_1) = f(x_2)$ then $x_1 = x_2$.  
  \newline
  \newline
  2. We say that a function $f : X \rightarrow Y$ is surjective or onto if, for every $y \in Y$ , there is at least one $x \in X$ such that $f(x) = y$.
  the range of f is the same as the codomain of f (range(f) = Y).
  \newline
  \newline
  3. We say that a function $f : X \rightarrow Y$ is bijective if $f$ is both injective and surjective (one-to-one and onto).   
  \newline
  \newline
  for every $y \in Y$ , there is exactly one $x \in X$ such that $f(x) = y$.
  \newline
  [example in word]
  \newpage
  \subsection{Composition of Functions}
    For functions $f : X \rightarrow Y$ and $g : Y \rightarrow Z$, the composite of f and g is the function $g \circ f : X \rightarrow Z$ defined by $(g \circ f)(x) = g(f(x))$ for all $x \in X$.
    \newline
    \newline
    The composite function $g \circ f$ exists whenever the range of f is a subset of the domain of g.
    \newline
    \newline
    In general, $g \circ f$ and $f \circ g$ are not the same composite functions. Associativity of composition (assuming they exist): $h \circ (g \circ f) = (h \circ g) \circ f$.  
\newline
\newline
\textbf{ Example: Take sets $X = \{\text{ all MATH1081 students }\}, Y = \{ 0, 1, \dots, 100 \}, Z = \{F, P, CR, D, HD\}.$  }
\newline
\newline 
Maps: $f = X \rightarrow Y; g = Y \rightarrow Z.$
\newline
\newline
A) $g \circ f: X \rightarrow Z.$
\newline
$(f \circ g) (y) = f(g(y)).$
\newline
[examples in word]
\newpage
\subsection{Identity and Inverse Functions}
\underline{ Identity Function:}
\newline
\newline
\boxed{i_x: x \rightarrow x; i_x(x) = x.}
\newline
\newline
For any function $f : X \rightarrow Y$ , we have $f \circ i_x = f = i_y \circ f$.   A function $g : Y \rightarrow X$ is an inverse of $f : X \rightarrow Y$ if
$g(f(x)) = x$ for all $x \in X$ and $f(g(y)) = y$ for all $y \in Y$,
\newline
or equivalently, $g \circ f = i_x$ and $f \circ g = i_y$ .
\newline
\newline
\boxed{\text{1. A function can have at most one inverse.}}
\newline
\newline
If $f : X \rightarrow Y$ has an inverse, then we say that $f$ is invertible, and we denote the inverse off by $f^{-1}$. Thus,$f^{-1} \circ f = i_x$ and $f \circ f^{-1} = i_y$.
\newline
\newline
If $g$ is the inverse of $f$, then $f$ is the inverse of $g$. Thus, $(f^{-1})^{-1} = f$.
\newline
\newline
[example in word]
\newpage
\subsection{Inverse Function Proofs}
\underline{ Theorem and Proof: }
\newline
\newline
\boxed{1. \text{ A function } f: X \rightarrow Y \text{ has at most 1 inverse} }
\newline
\newline
Proof:
\newline
\newline
$$\text{ Let } g_1, g_2: Y \rightarrow X \text{ be inverse of }f.$$
$$\text{ Then }: g_1 = g_i \circ i_y$$
$$ = g_i \circ (f \circ g_2)$$
$$ = (g_i \circ f) \circ g_2$$
$$ = i_x \circ g_2$$
$$ = g_2 \text{ End of proof }.$$
[example in word]
\newpage
\section{Number Theory and Relations}
\subsection{Numbers and Divisibility}
[topic 2 done in word (SteelsSlides1): maybe put in definitions here ?? that depends]
\newline
\newline
Number Set Notation:
\newline
\newline
1. The positive integers: $\mathbb{Z}^+ = \{ 1, 2, 3, \dots\},$
\newline
2. The natural numbers: $\mathbb{N} = \{ 0, 1, 2, 3, \dots\},$
\newline
3. The integers: $\mathbb{Z} = \{\dots, -2, -1, 0, 1, 2, \dots\},$
\newline
4. The rational numbers: $\mathbb{Q} = \{ m/n : m \in \mathbb{Z}, n \in \mathbb{Z}^+ \},$
\newline
5. The real numbers, $\mathbb{R}$ and the complex numbers $\mathbb{C}.$
\newline
\newline
Tests (Divisibility):
\newline
\newline
1. $2\text{ | } N$ if and only if the decimal expansion of N ends in an even integer
\newline
2. $5 \text{ | } N$ if and only if the last decimal digit of N is 5 or 0.
\newline
3. $3 \text{ | } N$ if and only if the sum of the decimal digits of N is divisible by 3.
\newline
3': $9 \text{ | } N$ if and only if the sum of the decimal digits of N is divisible by 9.
\newline
4. $11 \text{ | } N$if the alternating sum of the decimal digits of N is divisible by 11. 
\newline
(example: $1232 = 1 - 2 + 3 - 2 = 0$)
\newline
\newline
[proof in word]
\newpage
\subsection{Primes}
[in word]
\newline
\newline
\underline{Primes Definition: Formal:}
Another way of saying this is if $p$ is prime:
$$x \text{ | } p	\text{ implies }	x \in \{ -1,1,-p,p \}$$.
\newline
\newline
\underline{Theorems:}
\newline
\newline
1. If p is prime and $p | ab$, then $p | a$ or $p | b$,
\newline
2. If n is composite, then it has a prime factor less than or equal to $\sqrt[2]{n}$,
\newline
3. If no prime less than or equal to $\sqrt[2]{n}$ divides $n$ then $n$ is a prime,
\newline
4. Every integer $n \geq 2$ can be written uniquely as a product of a finite number of primes in increasing order i.e. $n = p_1^{m_1} * p_2^{m_2} \dots p_k^{m_k}$
for primes $p_1 < p_2 < \dots < p_k$ and exponents $m_1,m_2, \dots ,m_k \in \mathbb{Z}^+$.
\newline
\newline
Open Results about Primes:
\newline
\newline
1. A prime of the form $2^n + 1$ is called a Fermat prime.
\newline
2. A prime of the form $2^{n} - 1$ is called a Mersenne prime.
\newline
3. Two primes that differ by 2, are called twin primes.
For example, $3$ and $5$ are twin primes; so are $29$ and $31$.
\newline
4. The Goldbach Conjecture is that they are: it has been proved true for all numbers with fewer than about $17$ digits.
\newpage
\subsection{Common Divisors and Multiples}
[mostly on word]
\newline
All $a,b \in \mathbb{Z}$ have (at least) one common divisor, namely $1$, and so we can define the following:
\newline
For $a,b \in \mathbb{Z}$, not both zero, the positive integer d such that
$$1.	d \text{ | } a and d \text{ | } b,$$
$$2.	If c \text{ | } a and c \text{ | } b then c \le d.$$
is called the greatest common divisor of $a$ and $b$.
We write $d = gcd(a,b)$.
\newline
\newline
\boxed{ \text{ Begin by writing $a$ and $b$ as a product of primes. }}
\newline
\newline
\underline{ Properties of GCD:}
\newline
\newline
1. $\text{gcd }(a,b)$ is not affected by the signs of $a$ or $b$
\newline
2. Condition $(2)$ in the definition of gcd can be replaced by $(2')$ if $c \text{ | } a$ and $c \text{ | } b$ then $c \text{ | } d$.
\newline
3. For $a \in \mathbb{Z}^+, \text{gcd }(a,0) = a$.
\newline
\newline
\underline{Least Common Multiple}
\newline
\newline
All $a,b \in \mathbb{Z}$ have (at least) one common multiple, namely $ab$, and so we can define the following:
For $a,b \in \mathbb{Z}$, not both zero, the positive integer l such that
$$1)	a \text{ | } l \text{ and } b \text{ | } l$$
$$2)	\text{If } a \text{ | } c \text{ and } b \text{ | } c \text{ then } l \le c \text{ is called the least common multiple of } a \text{ and } b.$$ 
\newline
We write $l = \text{ lcm }(a,b)$.
\newline
\newline
Theorem:
\newline
\newline
\boxed{\text{For all positive integers $a$ and $b$; gcd}(a,b) x \text{ lcm }(a,b) = ab.}
\newpage
\subsection*{Quotient and Remainder}
\
[mostly in word]
\newline
\newline
\underline{The Quotient-Remainder Theorem (aka The Division Algorithm)}
\newline
\newline
If $a \in \mathbb{Z}$ and $b \in \mathbb{Z}^+$, then there exist unique $q,r \in Z$ such that (q: quotient; r: remainder):
$$a = bq + r	\text{ and }	0 \le r < b.$$
\newline
\newline
\boxed{\text{ Note: $q$ can be found using floor function; $q = \lfloor a / b \rfloor.$; then $r = a - qb$.}}
\newpage
\subsection{The Euclidean Algorithm}
\
[mostly in word]
\newline
\newline
\boxed{\text{If $a = bq + r$ then gcd$(a,b) =$ gcd$(b,r)$.}}
\newline
\newline
\underline{The Euclidian Algorithm: General Case [steps]}
\newline
\newline
$1)$	Let $a$ and $b$ be integers with $a > b \geq 0.$
\newline
$2)$	If $b = 0$, then gcd$(a,b) = a.$
\newline
$3)$	If $b > 0$, use the Quotient-Remainder theorem to write $a = bq + r$ where $0 \leq r < b$.	Then by our prevous result, gcd$(a,b) =$ gcd$(b,r).$
\newline
$4)$	Repeat steps 2 and 3 to find gcd$(b,r)$.
\newline
\newline
Example: Find gcd$(708, 540)$
$$708 = 540 \cdot 1 + 168,$$
$$540 = 168 \cdot 3 + 36,$$
$$168 = 36 \cdot 4 + 24,$$
$$36 = 24 \cdot 1 + 12,$$ 
$$24 = 12 \cdot 2 + 0.$$
So,
$$\text{gcd }(708,540) = 12.$$
\boxed{\text{Note: gcd is the last non-zero remainder.}}
\newline
\newline
\underline{Bezout's Identity}
\newline
\newline
For $a,b \in Z$ not both zero, there exist integers $x$ and $y$ (not unique) such that:
\newline
\newline
\boxed{ \text{gcd }(a,b) = ax + by.}
\newline
\newline
\boxed{\text{Theorem: Integers $a$ and $b$ are relatively prime if and only if there exists $x,y \in Z$ such that $ax + by = 1$.}}
\newline
\newline
\underline{Extended Euclidean Theorem:}
The Extended Euclidean Algorithm is a more efficient way of finding the numbers in Bézout's Identity: In looking for gcd$(a,b)$, assume $a > b > 0$.
\newline
1. We make up a table with five columns labelled $i, q_i, r_i, x_i, y_i$, where $i$ labels the rows.
\newline
2. We set row $1$ to be $1,0,a,1,0$ and row $2$ to be $2,0,b,0,1.$
Thus $q_1 = q_2 = 0; r_1 = a, r_2 = b; x_1 = y_2 = 1; x_2 = y_1 = 0.$
\newline
3. Then for $i$ from $3$ onwards, $q_i$ is the quotient on dividing $r_{i-2}$ by $r_{i-1}$ ($a$ divided by $b$ in the first case).
\newline
4. Then subtract $q_i$ times the rest of row $i - 1$ from row $i - 2$.
\newline
5. Repeat until we get $r_{n+1} = 0$ for some $n$, then stop.
Then the gcd is $r_n$ and $r_n = ax_n + by_n$, that is the last row before $r_i$ was zero gives the gcd, the $x$ and the $y$.
\newline
In fact a similar identity holds at each step: $r_i = ax_i + by_i$.
\newpage
\end{document}